MR-PFU: A suite of R packages for energy conversion chain analysis

### 2023-07-23 paper.Rmd
Working on the paper.

Historically, SEA has analyzed the ECCs of individual countries
using a large number of linked spreadsheets, 
with data in varying and inconsistent formats 
[@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa].
The authors of this paper and others in the field
wanted to expand SEA to cover all countries, but
the spreadsheet approach to data analysis was
deemed not scalable.
We created a suite of `R` packages to provide scalable SEA 
for every country in the world.
This paper describes the design of those packages and
demonstrates briefly their use.

@@

Several decisions influenced the design and APIs for the packages.
The decision involves data format. 
We selected the Physical Supply-Use Table (PSUT) framework,
a matrix approach to describing energy flows from 
extraction to
processing stages and, ultimately, 
to final demand [@Rocco:2016; @Guevara:2017; @Heun:2018].
The PSUT framework succinctly describes the flow of energy carriers
(``products'' in PSUT terminology) 
among energy conversion machines 
(``industries'' in PSUT terminology)
in a set of four matrices as described in the following table.

| Matrix      | rows x columns     | Name                           | Description                                              |
|:------------|:-------------------|:-------------------------------|:---------------------------------------------------------|
| **R**       | industry x product | Resource matrix                | Contains exogeneous energy inputs to an ECC              |
| **U**       | product x industry | Use matrix                     | Describes how each energy conversion device uses energy  |
| **U_EIOU**  | product x industry | Energy industry own use matrix | Describes how the energy industry uses energy            |
| **V**       | industry x product | Make matrix                    | Describes how each energy conversion device makes energy |
| **Y**       | product x industry | Final demand matrix            | Describes how each energy carrier is consumed            |

@@

| Package | Function |
|:--------|:---------|
| [RCLabels](https://matthewheun.github.io/RCLabels/)      | Manipulate row and column names in [matsindf](https://matthewheun.github.io/matsindf/) data frames |
| [matsbyname](https://matthewheun.github.io/matsbyname/)  | Perform matrix mathematics that respects row and column names |
| [matsindf](https://matthewheun.github.io/matsindf/)      | Store matrices in cells of a data frame, thereby enabling [tidyverse](https://www.tidyverse.org) syntax |

>>>>

Historically, SEA has analyzed the ECCs of individual countries, 
starting with primary and final stage data from the [IEA](https://www.iea.org/data-and-statistics/data-product/world-energy-balances)'s
world energy balances,
using a large number of linked spreadsheets, 
with data in varying and inconsistent formats 
[@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa].
The authors of this paper and others in the field
wanted to expand SEA to cover all countries, but
the spreadsheet approach to data analysis was
deemed not scalable.
We created a suite of `R` packages to provide scalable SEA 
for every country in the world.
This paper describes the design of those packages and
demonstrates briefly their use.

@@

Several decisions influenced the design and APIs for the packages.
The decision involves data format. 
We selected the Physical Supply-Use Table (PSUT) framework,
a matrix approach to describing energy flows from 
extraction to
processing stages and, ultimately, 
to final demand [@Rocco:2016; @Guevara:2017; @Heun:2018].
The PSUT framework succinctly describes the flow of energy carriers
("products" in PSUT terminology) 
among energy conversion machines 
("industries" in PSUT terminology)
in a set of six matrices as described in the following table.

| Matrix      | rows x columns     | Name                           | Description                                              |
|:------------|:-------------------|:-------------------------------|:---------------------------------------------------------|
| **R**       | industry x product | Resource matrix                | Contains exogeneous energy inputs to an ECC              |
| **U**       | product x industry | Use matrix                     | Describes how each energy conversion device uses energy  |
| **U_feed**  | product x industry | Feedstock use matrix           | Describes feedstock inputs to energy conversion devices  |
| **U_EIOU**  | product x industry | Energy industry own use matrix | Describes how the energy industry uses energy            |
| **V**       | industry x product | Make matrix                    | Describes how each energy conversion device makes energy |
| **Y**       | product x industry | Final demand matrix            | Describes how each energy carrier is consumed            |

@@

| Package | Function |
|:--------|:---------|
| [RCLabels](https://matthewheun.github.io/RCLabels/)      | Manipulates row and column names in [matsindf](https://matthewheun.github.io/matsindf/) data frames |
| [matsbyname](https://matthewheun.github.io/matsbyname/)  | Performs matrix mathematics that respects row and column names |
| [matsindf](https://matthewheun.github.io/matsindf/)      | Stores matrices in cells of a data frame, thereby enabling [tidyverse](https://www.tidyverse.org) syntax |
